Fast and optimal centroiding of faint stars

### 2015-03-24 ms/figure.tex
Update figure.tex

In this work, we focus on four centroiding methods. The first method
involves fitting a PSF profile, assuming that we have a good PSF estimate, to the star. 
The second method estimates the centroid of a star by fitting a 2d polynomial to
 the 3$\times$3 patch around the brightest pixel of the image convolved with the PSF. 
The third method centroids stars by
 smoothing the image of stars by a Gaussian kernel of a fixed size,
 and then applying the same 3$\times$3 polynomial trick to the smooth
 image. This method is fast and does not require any knowledge of the 
PSF. The last method we consider, is the center of light centroiding applied to the
 7$\times$7 patch around the brightest pixel of the image.

@@

Then, we fit a simple 2d polynomial 
$P(x,y)=a+bx+cy+dx^2+exy+fy^2$ 
to the 3$\times$3 patch around the brightest pixel of the convolved image $Y^{s}$.
Upon constructing a universal 9$\times$6 design matrix
\begin{equation}
    \mathbf{A} = 
    \begin{bmatrix}
        1 & x_{1} & y_{1} & x_{1}^{2} & x_{1}y_{1} & y_{1}^{2} \\
        . & . & . & . & . & .  \\
        . & . & . & . & . & .  \\
        . & . & . & . & . & .  \\
        1 & x_{9} & y_{9} & x_{9}^{2} & x_{9}y_{9} & y_{9}^{2}
    \end{bmatrix},
\end{equation}
the free parameters $\{a,b,c,d,e,f\}$
(hereafter compactly denoted by $\mathbf{X}$) can be determined by 
\beq
\mathbf{X} = (\mathbf{A}^{T}\mathbf{A})^{-1}\mathbf{A}^{T}\mathbf{Z},
\label{linearfit}
\eeq
where $\mathbf{Z}$ is given by $(z_{1},...,z_{9})^{T}$,
with $z_{i}$, being the brightness of the $i-$th pixel of the 3$\times$3 patch centered on the brightest pixel of $Y^{(s)}$.
Afterwards, the best fit parameters can be used to compute the centroid coordinate

@@

The procedure of convolving the image of star with the PSF results in a
more well-sampled image, which eventually, leads to having more
information regarding the centroids of stars in the 3$\times$3 patch around the
brightest pixel. In addition, a simple 2d polynomial will provide a better fit 
since convolution with the PSF makes the variation of the brightness of the image
 across the 3$\times$3 patch very smooth.

\item{\bf Default polynomial centroiding}\quad In the case that 
we do not have know the PSF at the position of star, we change 
the smoothing step in the following way. Instead of smoothing the image 
by convolving it with the PSF, smoothing is done by convolving the image 
with a Gaussian kernel of a fixed size 
\beq
k(\mathbf{x}) = \frac{1}{2\pi w^2}\exp(-\mathbf{x}^{2}/2w^{2}),
\eeq
where throughout this study, the full-width at half-maximum of the Gaussian kernel is held at
a fixed value of 2.8 pixels (corresponding to w $\simeq$ 1.2 pixels). Convolution is done as follows
\begin{eqnarray}
Y^{(s)} &=& Y \star \mathcal{K}, \\
Y^{(s)}_{[i,j]} &=& \sum_{k,l}Y_{[i-k,j-l]}\mathcal{K}_{[k,l]},
\end{eqnarray}
where $Y$ is the image of the star, $Y^{(s)}$ is the smooth image, and $\mathcal{K}$ is a 7$\times$7 
array whose elements are given by the Gaussian kernel
\beq
\mathcal{K}_{[k,l]} = k(x_{k},y_{l}).
\eeq 
Then we apply the same 2d polynomial method to the 3$\times$3 patch centered on the brightest
pixel of the smooth image $Y^{(s)}$. Therefore, for a given star and a smoothing kernel,
the outcome of equation (\ref{linearfit}) can be
plugged into equation (\ref{center}) to find the centroid estimate
of the star. This is inspired by the 3$\times$3 quartic approximation 
used in the \textsl{Sloan Digital Sky Surveys} photometric pipeline \citep{sdss}.

>>>>

In this work, we focus on four centroiding methods. The first method
involves fitting a PSF profile, assuming that we have a good PSF estimate, to the star. 
The second method estimates the centroid of a star by fitting a 2d second-order polynomial to
 the 3$\times$3 patch around the brightest pixel of the image convolved with the PSF. 
The third method centroids stars by
 smoothing the image of stars by a Gaussian kernel of a fixed size,
 and then applying the same 3$\times$3 polynomial trick to the smooth
 image. This method is fast and does not require any knowledge of the 
PSF. The last method we consider, is the center of light centroiding applied to the
 7$\times$7 patch around the brightest pixel of the image.

@@

Then, we fit a simple 2d second-order polynomial 
$P(x,y)=a+bx+cy+dx^2+exy+fy^2$ 
to the 3$\times$3 patch around the brightest pixel of the convolved image $Y^{s}$.
Upon constructing a universal 9$\times$6 design matrix
\begin{equation}
    \mathbf{A} = 
    \begin{bmatrix}
        1 & x_{1} & y_{1} & x_{1}^{2} & x_{1}y_{1} & y_{1}^{2} \\
        . & . & . & . & . & .  \\
        . & . & . & . & . & .  \\
        . & . & . & . & . & .  \\
        1 & x_{9} & y_{9} & x_{9}^{2} & x_{9}y_{9} & y_{9}^{2}
    \end{bmatrix},
\end{equation}
the free parameters $\{a,b,c,d,e,f\}$
(hereafter compactly denoted by $\mathbf{X}$) can be determined by 
\beq
\mathbf{X} = (\mathbf{A}^{T}\mathbf{A})^{-1}\mathbf{A}^{T}\mathbf{Z},
\label{linearfit}
\eeq
where $\mathbf{Z}$ is given by $(z_{1},...,z_{9})^{T}$,
with $z_{i}$, being the brightness of the $i-$th pixel of the 3$\times$3 patch centered on the brightest pixel of $Y^{(s)}$.
Afterwards, the best fit parameters can be used to compute the centroid coordinate

@@

The procedure of convolving the image of star with the PSF results in a
more well-sampled image, which eventually, leads to having more
information regarding the centroids of stars in the 3$\times$3 patch around the
brightest pixel. In addition, a simple second-order polynomial will provide a better fit 
since convolution with the PSF makes the variation of the brightness of the image
 across the 3$\times$3 patch very smooth.

\item{\bf Default polynomial centroiding}\quad In the case that 
we do not have know the PSF at the position of star, we change 
the smoothing step in the following way. Instead of smoothing the image 
by convolving it with the PSF, smoothing is done by convolving the image 
with a Gaussian kernel of a fixed size 
\beq
k(\mathbf{x}) = \frac{1}{2\pi w^2}\exp(-\mathbf{x}^{2}/2w^{2}),
\eeq
where throughout this study, the full-width at half-maximum of the Gaussian kernel is held at
a fixed value of 2.8 pixels (corresponding to w $\simeq$ 1.2 pixels). Convolution is done as follows
\begin{eqnarray}
Y^{(s)} &=& Y \star \mathcal{K}, \\
Y^{(s)}_{[i,j]} &=& \sum_{k,l}Y_{[i-k,j-l]}\mathcal{K}_{[k,l]},
\end{eqnarray}
where $Y$ is the image of the star, $Y^{(s)}$ is the smooth image, and $\mathcal{K}$ is a 7$\times$7 
array whose elements are given by the Gaussian kernel
\beq
\mathcal{K}_{[k,l]} = k(x_{k},y_{l}).
\eeq 
Then we apply the same 2d second-order polynomial method to the 3$\times$3 patch centered on the brightest
pixel of the smooth image $Y^{(s)}$. Therefore, for a given star and a smoothing kernel,
the outcome of equation (\ref{linearfit}) can be
plugged into equation (\ref{center}) to find the centroid estimate
of the star. This is inspired by the 3$\times$3 quartic approximation 
used in the \textsl{Sloan Digital Sky Surveys} photometric pipeline \citep{sdss}.
